---
title: "Distributions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Distributions}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

These are the currently implemented distributions. 

| Name                                | univariateML function  | Package    | Parameters               | Support        | 
| ----------------------------------- | ---------------------- | ---------- | ------------------------ | -------------- | 
| Cauchy distribution                 | `mlcauchy`             | stats      | `location`,`scale`       | $\mathbb{R}$   | 
| Gumbel distribution                 | `mlgumbel`             | extraDistr | `mu`, `sigma`            | $\mathbb{R}$   | 
| Laplace distribution                | `mllaplace`            | extraDistr | `mu`, `sigma`            | $\mathbb{R}$   | 
| Logistic distribution               | `mllogis`              | stats      | `location`,`scale`       | $\mathbb{R}$   | 
| Normal distribution                 | `mlnorm`               | stats      | `mean`, `sd`             | $\mathbb{R}$   | 
| Student t distribution              | `mlstd`                | fGarch     | `mean`, `sd`, `nu`       | $\mathbb{R}$   | 
| Generalized Error distribution      | `mlged`                | fGarch     | `mean`, `sd`, `nu`       | $\mathbb{R}$   | 
| Skew Normal distribution            | `mlsnorm`              | fGarch     | `mean`, `sd`, `xi`       | $\mathbb{R}$   | 
| Skew Student t distribution         | `mlsstd`               | fGarch     | `mean`, `sd`, `nu`, `xi` | $\mathbb{R}$   | 
| Skew Generalized Error distribution | `mlsged`               | fGarch     | `mean`, `sd`, `nu`, `xi` | $\mathbb{R}$   | 
| Beta prime distribution             | `mlbetapr`             | extraDistr | `shape1`, `shape2`       | $(0, \infty)$  |
| Exponential distribution            | `mlexp`                | stats      | `rate`                   | $[0, \infty)$  | 
| Gamma distribution                  | `mlgamma`              | stats      | `shape`,`rate`           | $(0, \infty)$  | 
| Inverse gamma distribution          | `mlinvgamma`           | extraDistr | `alpha`, `beta`          | $(0, \infty)$  |
| Inverse Gaussian distribution       | `mlinvgauss`           | actuar     | `mean`, `shape`          | $(0, \infty)$  |
| Inverse Weibull distribution        | `mlinvweibull`         | actuar     | `shape`, `rate`          | $(0, \infty)$  |
| Log-logistic distribution           | `mlllogis`             | actuar     | `shape`, `rate`          | $(0, \infty)$  |
| Log-normal distribution             | `mllnorm`              | stats      | `meanlog`, `sdlog`       | $(0, \infty)$  |
| Lomax distribution                  | `mllomax`              | extraDistr | `lambda`, `kappa`        | $[0, \infty)$  | 
| Rayleigh distribution               | `mlrayleigh`           | extraDistr | `sigma`                  | $[0, \infty)$  | 
| Weibull distribution                | `mlweibull`            | stats      | `shape`,`scale`          | $(0, \infty)$  |
| Log-gamma distribution              | `mllgamma`             | actuar     | `shapelog`, `ratelog`    | $(1, \infty)$  |
| Pareto distribution                 | `mlpareto`             | extraDistr | `a`, `b`                 | $[b, \infty)$  |
| Beta distribution                   | `mlbeta`               | stats      | `shape1`,`shape2`        | $(0, 1)$       |
| Kumaraswamy distribution            | `mlkumar`              | extraDistr | `a`, `b`                 | $(0, 1)$       |
| Logit-normal                        | `mllogitnorm`          | logitnorm  | `mu`, `sigma`            | $(0, 1)$       |
| Uniform distribution                | `mlunif`               | stats      | `min`, `max`             | $[\min, \max]$ |
| Power distribution                  | `mlpower`              | extraDistr | `alpha`, `beta`          | $[0, a)$       |


This package follows a naming convention for the `ml***` functions. To access the
documentation of the distribution associated with an `ml***` function, write `package::d***`. 
For instance, to find the documentation for the log-gamma distribution write

```{r docs, eval = FALSE}
?actuar::dlgamma
```

## Problematic Distributions
### Lomax Distribution
The maximum likelihood estimator of the [Lomax distribution](https://en.wikipedia.org/wiki/Lomax_distribution) frequently fails to 
exist. For assume $\kappa\to\lambda^{-1}\overline{x}^{-1}$ and $\lambda\to0$. The density $\lambda\kappa\left(1+\lambda x\right)^{-\left(\kappa+1\right)}$
is approximately equal to $\lambda\kappa\left(1+\lambda x\right)^{-\left(\lambda^{-1}\overline{x}^{-1}+1\right)}$
when $\lambda$ is small enough. Since $\lambda\kappa\left(1+\lambda x\right)^{-\left(\lambda^{-1}\overline{x}^{-1}+1\right)}\to\overline{x}^{-1}e^{-\overline{x}^{-1}x}$,
the density converges to an exponential density. 

```{r lomax, fig.width = 6, fig.height = 5}
eps <- 0.1
x <- seq(0, 3, length.out = 100)
plot(dexp, 0, 3, xlab = "x", ylab = "Density", main = "Exponential and Lomax")
lines(x, extraDistr::dlomax(x, lambda = eps, kappa = 1 / eps), col = "red")
```
